# Occurrences

#### Load the required packages with:

`` `{r warning=FALSE, error=FALSE, message=FALSE, results='hide'}

```{r}
library("devtools")
library("rgbif")
library("fpc")
library("httr")
library("stringr")
library("XML")
library("RCurl")
library("sp")
library("rgeos") 
library("maps") 
library("RColorBrewer")
#library("mapdata") 
library("rOpenSearch")
library("rgbif")
library("knitr")
library("rworldmap")
```

#### Getting the White Shark occurrences from GBIF wit rgbif

```{r}
# get the occurrences from GBIF using rgbif
key <- name_backbone(name='Carcharodon carcharias', kingdom='animalia')$speciesKey
occurrences <- occ_search(taxonKey=key, limit=1000, return='data', hasCoordinate=TRUE)
occurrences <- occurrences[complete.cases(occurrences),]
```

This create the data frame occ with `r nrow(occurrences)` entries. The first five entries look like:

```{r results='asis', echo=FALSE}
kable(head(occurrences, 5), format = "html")
```

The occurrences can be plotted to create the map:

```{r results='asis', message=FALSE, echo=FALSE}
gbifmap(occurrences)
```

Apply the DBSCAN clustering algorithm to identify representative occurence clusters.

```{r}
occurrences <- cbind(occurrences$decimalLongitude, occurrences$decimalLatitude)

eps <- 5
minpts <- 20
dbscan.res <- dbscan(occurrences, eps=eps, MinPts=minpts)
```

The DBSCAN algorithm identified `r max(dbscan.res$cluster)`.

The cluster can be plotted using the cluster id color:

```{r}
plot(occurrences, col=dbscan.res$cluster)
```

```{r}
 mbr <- list()
  
  #extract the minimum bounding box for each cluster
  for(i in 1:max(dbscan.res$cluster)) {

    # get the cluster
    cl <- (occurrences[dbscan.res$cluster %in% i,])
  
    # create the matrix with the cluster minimum bounding box
    coords <- matrix(nrow=5, ncol=2, byrow=TRUE, data=c(
      min(cl[,1]), min(cl[,2]),
      max(cl[,1]), min(cl[,2]),
      max(cl[,1]), max(cl[,2]),
      min(cl[,1]), max(cl[,2]),
      min(cl[,1]), min(cl[,2])))
  
    # get the cluster geospatial envelope
    if (gArea(gEnvelope(SpatialPoints(coords)))>0) mbr[[length(mbr)+1]] <- gEnvelope(SpatialPoints(coords))

  }
  print(mbr)
```
